---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# vsp

The goal of vsp is to ...

## Installation

`vsp` is experimental and nascent. If you are a collaborator on the project, you can install the development version with:

``` r
install.packages("devtools")
devtools::install_github("alexpghayes/vsp", auth_token = NEED_TO_DO_THIS)
```

You'll need to set up a Github auth token. See the documentation in `?remotes::install_github()` for details on how to do this.

## Example

Fitting a `vsp` factor analysis:

```{r}
library(vsp)
library(Matrix)

# make a random sparse graph
M <- rsparsematrix(12, 12, nnz = 40)
M2 <- sign(M)^2

# vintage sparse pca
fa <- vsp(M2, k = 7)
fa
```

The screeplot:

```{r}
screeplot(fa)
```

```{r}
project_pca(fa)
```

```{r}
project_varimax(fa)
```

```{r}
pairs(project_pca(fa))
```

```{r}
pairs(project_varimax(fa))
```

think about how else to visualize these matrices. distributions of loadings? am i actually plotting the loadings here or did i just do some nonsense?

```{r}
plot_simulation_test(M2, k = 3)
```


```{r eval = FALSE}
library(tidyverse)

fa$B %>%
  as.matrix() %>%
  as_tibble() %>%
  mutate(row = row_number()) %>%
  gather(k, v, -row) %>%
  ggplot(aes(row, k, fill = v)) +
  geom_tile() +
  scale_fill_viridis_c()

image(fa$B)
image(fa$Z)
```
