---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(ggplot)
theme_set(theme_minimal())
```

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- badges: end -->

**Disclaimer:** This package is a work in progress. Some basic functionality is available, but the API is subject to change. Code to reproduce the results in *Vintage Factor Analysis with Varimax Performs Statistical Inference* is [available here](https://github.com/RoheLab/vsp-paper).

# vsp

The goal of `vsp` to enable fast, spectral estimation of latent factors in random dot product graphs. Under mild assumptions, the `vsp` estimator is consistent for (degree-corrected) stochastic blockmodels, (degree-corrected) mixed-membership stochastic blockmodels, degree-corrected overlapping stochastic blockmodels.

More generally, the `vsp` estimator is consistent for random dot product graphs that can be written in the form

```
E(A) = Z B Y^T
```

where `Z` and `Y` satisfy the varimax assumptions of [1]. `vsp` works on directed and undirected graphs, and on weighted and unweighted graphs.

## Installation

You can install the development version of `vsp` with:

``` r
install.packages("devtools")
devtools::install_github("RoheLab/vsp")
```

## Example

Obtaining estimates from `vsp` is straightforward. We recommend representing networks as [`igraph`](https://igraph.org/r/) objects or (sparse) `Matrix` objects from the [`Matrix`](https://cran.r-project.org/web/packages/Matrix/index.html) package (see also `?Matrix`).

Once you have your network in one of these formats, you can get estimates by calling the `vsp()` function. The result is an object of class `"vsp"`.

```{r}
library(vsp)
library(Matrix)

# create a random sparse adjacency matrix
M <- sign(rsparsematrix(12, 12, nnz = 40))^2

# vintage sparse pca
fa <- vsp(M, k = 5)
fa
```

`vsp()` is an S3 generic, so can similarly just pass an `igraph` object. This means that `vsp` also supports [`tidygraph`](https://github.com/thomasp85/tidygraph), which is built on top of `igraph`. We use `enron` network from `igraphdata` package to demonstrate this functionality.

```{r}
library(igraph)
data(enron, package = "igraphdata")

enron
```

```{r}
enron_fa <- vsp(enron, k = 30, center = FALSE)
enron_fa
```

The default print method has a large amount of valuable information. First, we learn that the graph has 184 nodes, and that we have used `k = 30` as our guess of the rank of `A`, the network adjacency matrix. The fifth singular value of `A` is 0.2077. This is of interest, as we expect `lambda[k] ~ 0` to indicate that we have we have overestimated the rank of `A`, in say, blockmodels. 

The pre-processing section indicates that the matrix representation of the graph was row- and column-centered before estimation (`center = TRUE`), and that we used the graph Laplacian rather than the adjacency matrix (`normalize = TRUE`).

By default, `normalize = TRUE` not only uses the graph Laplacian, but also regularizes the graph Laplacian using mean in-degree and out-degree. In particular, we form the graph Laplacian as follows:

```
# TODO
```

In general, you do not need to tune `tau_row` and `tau_col`. To turn off regularization (not recommended!), see `tau_row = 0` and `tau_col = 0`. This often leads to eigenvector localization and thus overfitting.

```{r}
vsp(M, k = 5, tau_row = 0, tau_col = 0)
```

The components `U`, `d`, and `V` are the standard elements of the singular value decomposition of your network. In particular, you can approximately reconstruct the matrix representation of your network as follows:

```{r}
fa2 <- vsp(M, k = 5, center = FALSE, normalize = FALSE)

fa2$U %*% diag(fa2$d) %*% t(fa2$V)
```

Note that you don't want to do this for large networks since the reconstruction will be dense. Similarly, the varimax-rotated network approximate reconstruction is

```{r}
fa$Z %*% fa$B %*% t(fa$Y)
```

If you want to directly manipulate the estimates, you can access them as we have above with `fa$B` and so on and so forth. Note that these are `matrix` and `Matrix` objects, and you may need to perform type coercions. We also provide some utility functions to extract estimates into `tibbles`:

```{r}
get_varimax_z(fa)
```

```{r}
get_varimax_y(fa)
```

To visualize a screeplot of the singular value, use:

```{r}
screeplot(fa)
```

At the moment, we also enjoy using pairs plots of the factors as a diagnostic measure:

```{r}
plot_varimax_z_pairs(fa, 1:3)
```

```{r}
plot_varimax_y_pairs(fa, 1:3)
```

Similarly, an IPR pairs plot can be a good way to check for singular vector localization (and thus overfitting!).

```{r}
# TODO
```

TODO: visualizing B

TODO: selecting k


## Tidygraph integration

```{r}
library(tidygraph)

sbm <- play_blocks(25, 25, 0.1)
```

`vsp` also works on `igraph` and `tidygraph` objects:

```{r}
fa2 <- vsp(sbm, k = 5)
```

You can then add the resulting factor loadings back into the node information about the graph:

```{r}
# if you already have a vsp object
sbm_fa2 <- sbm %>% 
  bind_varimax_z(fa2)

sbm_fa2
```

## References

1. Rohe, K. & Zeng, M. *Vintage Factor Analysis with Varimax Performs Statistical Inference*. 2020+.

