---
title: "vsp"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vsp}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Obtaining estimates from `vsp` is straightforward. We recommend representing networks as [`igraph`](https://igraph.org/r/) objects or (sparse) `Matrix` objects from the [`Matrix`](https://cran.r-project.org/web/packages/Matrix/index.html) package (see also `?Matrix`).

Once you have your network in one of these formats, you can get estimates by calling the `vsp()` function. The result is an object of class `"vsp_fa"`.

```{r}
library(vsp)
library(Matrix)

# create a random sparse adjacency matrix
M <- sign(rsparsematrix(12, 12, nnz = 40))^2

# vintage sparse pca
fa <- vsp(M, rank = 5, scale = TRUE, center = TRUE, recenter = TRUE)
fa
```

Here we demonstrate `vsp` usage on an `igraph` object, using the `enron` network from `igraphdata` package to demonstrate this functionality. First we peak at the graph:

```{r}
library(igraph)
data(enron, package = "igraphdata")

enron
```

Now we estimate:

```{r}
library(Matrix)

enron_fa <- vsp(enron, rank = 30)
enron_fa
```

The default print method has a large amount of valuable information. First, we learn that the graph has 184 nodes, and that we have used `rank = 30` as our guess of the rank of `A`, the network adjacency matrix. The 30th singular value of `A` is 0.2039. This is of interest, as we expect `lambda[k] ~ 0` to indicate that we have we have overestimated the rank of `A`, in say, blockmodels. 

The pre-processing section indicates that the matrix representation of the graph was *not* row- and column-centered before estimation (`center = FALSE`), and that we used the graph Laplacian rather than the adjacency matrix (`normalize = TRUE`).

By default, `normalize = TRUE` not only uses the graph Laplacian, but also regularizes the graph Laplacian using mean in-degree and out-degree. See [1] for details on how we form the regularized graph Laplacian.

In general, you do not need to tune `tau_row` and `tau_col`. To turn off regularization (not recommended!), see `tau_row = 0` and `tau_col = 0`. This often leads to eigenvector localization and thus overfitting.

```{r}
vsp(M, rank = 5, tau_row = 0, tau_col = 0)
```

The components `U`, `d`, and `V` are the standard elements of the singular value decomposition of your network.

At the moment, we are performing centering explicitly, which is inefficient and converts adjacency matrices from sparse into dense matrices. Setting `center = TRUE` produces a warning to remind you of this. We plan to implement implicit centering shortly.

```{r warning = TRUE}
vsp(M, rank = 5, center = TRUE)
```

You can approximately reconstruct the matrix representation of your network as follows:

```{r}
fa2 <- vsp(M, rank = 5, normalize = FALSE)

fa2$u %*% diag(fa2$d) %*% t(fa2$v)
```

In the example above, this will *not* approximate `M`, but rather the regularized graph Laplacian formed from `M`. Similarly, the varimax-rotated network approximate reconstruction is

```{r}
fa$Z %*% fa$B %*% t(fa$Y)
```

In general, you won't want to explicit reconstruct the network matrix for large networks since the reconstruction will be dense. 

If you want to directly manipulate the estimates, you can access them as we have above with `fa$B` and so on and so forth. Note that these are `matrix` and `Matrix` objects, and you may need to perform type coercions. We also provide some utility functions to extract estimates into `tibbles`:

```{r}
get_varimax_z(fa)
```

```{r}
get_varimax_y(fa)
```

To visualize a screeplot of the singular value, use:

```{r}
# screeplot(fa)
```

At the moment, we also enjoy using pairs plots of the factors as a diagnostic measure:

```{r}
# plot_varimax_z_pairs(fa, 1:3)
```

```{r}
# plot_varimax_y_pairs(fa, 1:3)
```

Similarly, an IPR pairs plot can be a good way to check for singular vector localization (and thus overfitting!).

```{r}
# TODO
```

TODO: visualizing B

TODO: selecting k


## Tidygraph integration

```{r}
library(tidygraph)

sbm <- play_blocks(25, 25, 0.1)
```

`vsp` also works on `igraph` and `tidygraph` objects:

```{r}
fa2 <- vsp(sbm, rank = 5)
```

You can then add the resulting factor loadings back into the node information about the graph:

```{r}
# if you already have a vsp object
sbm_fa2 <- sbm %>% 
  bind_varimax_z(fa2)

sbm_fa2
```